Concurrent Programming in Harmony

### 2021-07-11 book/paper.tex
Latest

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{empty()} & returns an empty bag\\
\hline
\texttt{fromSet($s$)} & create a bag from set $s$\\
\hline
\texttt{fromList($t$)} & convert list $t$ into a bag \\
\hline
\texttt{count($b$, $e$)} & count how many times $e$ occurs in bag $b$\\
\hline
\texttt{bchoose($b$)} & like \texttt{choose(s)}, but applied to a bag\\
\hline
\texttt{add}(\textit{pb}, $e$) & add one copy of $e$ to bag \texttt{!}\textit{pb}\\
\hline
\texttt{remove}(\textit{pb}, $e$) & remove one copy of $e$ from bag \texttt{!}\textit{pb}\\
\hline
\end{tabular}

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{issubset($s$, $t$)} & returns whether $s$ is a subset of $t$ \\
\hline
\texttt{issuperset($s$, $t$)} & returns whether $s$ is a superset of $t$ \\
\hline
\end{tabular}

@@

\newpage
Clarifications:
\begin{itemize}
\item The \texttt{Address} instruction expects two values on the stack.
The top value must be an address value, representing a dictionary 
The other value must be a key into the dictionary.
The instruction then computes the address of the given key.
\item Even though Harmony code does not allow taking addresses of thread variables, both
shared and thread variables can have addresses.
\item The \texttt{Load}, \texttt{LoadVar}, \texttt{Del}, \texttt{DelVar},
and \texttt{Stop} instructions have an optional
variable name: if omitted the top of the stack must contain the address of
the variable.
\item \texttt{Store} and \texttt{StoreVar} instructions have an optional
variable name.  In both cases the value to be assigned is on the top
of the stack.  If the name is omitted, the address is underneath that
value on the stack.
\item The effect of the \texttt{Apply} instructions depends much on $m$.
If $m$ is a dictionary, then \texttt{Apply} finds $i$ in the dictionary
and pushes the value.
If $m$ is a program counter, then \texttt{Apply} invokes method $m$ by
pushing the current program counter and setting the program counter to
$m$.  $m$ is supposed to leave the result on the stack.
\item The \texttt{Frame} instruction pushes the value of the thread
register (\emph{i.e.}, the values of the thread variables) onto the
stack.  It initializes the \texttt{result} variable to the empty dictionary.
The \texttt{Return} instruction restores the thread register by popping
its value of the stack.
\item All method calls have exactly one argument, although it sometimes
appears otherwise:
\begin{itemize}
\item \texttt{m()} invokes method \texttt{m} with the empty dictionary \texttt{()} as argument;
\item \texttt{m($a$)} invokes method \texttt{m} with argument $a$;
\item \texttt{m($a$, $b$, $c$)} invokes method \texttt{m} with tuple \texttt{($a$, $b$, $c$)} as argument.
\end{itemize}
The \texttt{Frame} instruction unpacks the argument to the method and places them into thread variables by the given names.
\item Every \texttt{Stop} instruction must immediately be followed by a 
\texttt{Continue} instruction.
\end{itemize}

@@

Local variables all declared.  They can be declared in
\texttt{def} statements (i.e., arguments),
\texttt{let} statements, and in \texttt{for} loops.
Also, each method has an implicitly declared
\textit{result} variable.
Local variables are tightly scoped and cannot be shared
between threads.
While in theory one method can be declared within another,
they cannot share variables either.
All other variables are global and must be initialized
before any threads are spawned.

@@

>>>>

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{empty()} & returns an empty bag\\
\hline
\texttt{fromSet($s$)} & create a bag from set $s$\\
\hline
\texttt{fromList($t$)} & convert list $t$ into a bag \\
\hline
\texttt{count($b$, $e$)} & count how many times $e$ occurs in bag $b$\\
\hline
\texttt{bchoose($b$)} & like \texttt{choose(s)}, but applied to a bag\\
\hline
\texttt{add}($b$, $e$) & add one copy of $e$ to bag $b$ \\
\hline
\texttt{remove}($b$, $e$) & remove one copy of $e$ from bag $b$ \\
\hline
\texttt{combinations}($b$, $k$) & return set of all \emph{subbags} of size $k$ \\
\hline
\end{tabular}

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
\texttt{issubset($s$, $t$)} & returns whether $s$ is a subset of $t$ \\
\hline
\texttt{issuperset($s$, $t$)} & returns whether $s$ is a superset of $t$ \\
\hline
\texttt{add}($s$, $e$) & returns $s \cup \{ e \}$ \\
\hline
\texttt{remove}($s$, $e$) & returns $s - \{ e \}$ \\
\hline
\texttt{combinations}($b$, $k$) & returns set of all subsets of size $k$ \\
\hline
\end{tabular}

@@

\newpage
Clarifications:
\begin{itemize}
\item The \texttt{Address} instruction expects two values on the stack.
The top value must be an address value, representing a dictionary 
The other value must be a key into the dictionary.
The instruction then computes the address of the given key.
\item Even though Harmony code does not allow taking addresses of thread variables, both
shared and thread variables can have addresses.
\item The \texttt{Load}, \texttt{LoadVar}, \texttt{Del}, \texttt{DelVar},
and \texttt{Stop} instructions have an optional
variable name: if omitted the top of the stack must contain the address of
the variable.
\item \texttt{Store} and \texttt{StoreVar} instructions have an optional
variable name.  In both cases the value to be assigned is on the top
of the stack.  If the name is omitted, the address is underneath that
value on the stack.
\item The effect of the \texttt{Apply} instructions depends much on $m$.
If $m$ is a dictionary, then \texttt{Apply} finds $i$ in the dictionary
and pushes the value.
If $m$ is a program counter, then \texttt{Apply} invokes method $m$ by
pushing the current program counter and setting the program counter to
$m$.  $m$ is supposed to leave the result on the stack.
\item The \texttt{Frame} instruction pushes the value of the thread
register (\emph{i.e.}, the values of the thread variables) onto the
stack.  It initializes the \texttt{result} variable to \texttt{None}.
The \texttt{Return} instruction restores the thread register by popping
its value of the stack.
\item All method calls have exactly one argument, although it sometimes
appears otherwise:
\begin{itemize}
\item \texttt{m()} invokes method \texttt{m} with the empty dictionary \texttt{()} as argument;
\item \texttt{m($a$)} invokes method \texttt{m} with argument $a$;
\item \texttt{m($a$, $b$, $c$)} invokes method \texttt{m} with tuple \texttt{($a$, $b$, $c$)} as argument.
\end{itemize}
The \texttt{Frame} instruction unpacks the argument to the method and places them into thread variables by the given names.
\item Every \texttt{Stop} instruction must immediately be followed by a 
\texttt{Continue} instruction.
\end{itemize}

@@

Local variables all declared.  They can be declared in
\texttt{def} statements (i.e., arguments),
\texttt{let} statements, and in \texttt{for} loops.
Also, each method has an implicitly declared
\textit{result} variable, which is initialized to \texttt{None}.
Local variables are tightly scoped and cannot be shared
between threads.
While in theory one method can be declared within another,
they cannot share variables either.
All other variables are global and must be initialized
before any threads are spawned.

@@

In case of dictionaries, you can access both keys and values:
\begin{code}
\texttt{for} $k:v$ in \{ .a: 1, .b: 2 \}: ...
\end{code}
